---
title: "R Notebook"
output: html_notebook
---

```{r}
library(scales)
library(olsrr)
library(magrittr)
library(dplyr) 
library(mlbench)
library(tidyverse)
library(caret)
library(corrplot)



```



##loading and cleaning
```{r}
baseball = read.csv("/Users/taha/Downloads/baseball.csv")
colSums(is.na(baseball))
baseball <-  na.omit(baseball)

```




##looking into p values and correlations
```{r}

# picking lowest p vlaues
fit_for_pvalues = lm(W~.,baseball)
anova(fit_for_pvalues)


# lowest p values =>  RankSeason, OOBP, RA, RS, Year, SLG

# high correlations
# numeric_vars <- baseball %>%
#                   select(-Team,-League)
# 
# correlation_matrix <- cor(numeric_vars)
# correlation_matrix
# highlyCorrelated <- findCorrelation(correlation_matrix,cutoff=0.7)
# colnames(numeric_vars[highlyCorrelated])
# # [1] "RS"  "SLG" "OBP"
# corrplot(correlation_matrix, method="circle")



```





##complete-reduced partial F test
```{r}
data <- subset(baseball,select = -c(Team,RankPlayoffs,RankSeason))

# reduced model
model_r <- lm(W~ OOBP+ RA+ RS+ Year, data=data )
summary(model_r)
anova(model_r)


# complete model
model_c <- lm(W~., data=data )
summary(model_c)
anova(model_c)


dfE_complete = 101
dfE_reduced = 107
df_drop = dfE_reduced - dfE_complete
SSEreduced = 634.79
SSEcomplete = 595.16


MSEcomplete = SSEcomplete / dfE_complete        #5.892673
MSdrop = (SSEreduced - SSEcomplete) /df_drop    # 6.605

F_ts = MSdrop / MSEcomplete                     #1.120883

p_value = pf(F_ts,df_drop,dfE_complete,lower.tail=F) #0.355599 ????



```



##plots for checking assumptions
```{r, echo=FALSE,results='hide',fig.keep='all'}


fit.lm <- model_r
fit.stdres = rstandard(fit.lm)


plot.new()
qqnorm(fit.stdres, ylab="Standardized Residuals", xlab="Normal Scores",  main="qqplot of residuals") 
qqline(fit.stdres)

plot(seq(1,nrow(data)),fit.stdres,main="Time plot")
abline(0, 0)

plot(fit.lm)

# reduced model variables RankPlayoffs+ RankSeason+ OOBP+ RA+ RS+ Year

plot(data$OOBP,fit.stdres,ylab="Residuals", xlab="OOBP" )
abline(0, 0) 

plot(data$RA,fit.stdres,ylab="Residuals", xlab="RA")
abline(0, 0)


plot(data$RS,fit.stdres,ylab="Residuals", xlab="RS")
abline(0, 0)

plot(data$Year,fit.stdres,ylab="Residuals", xlab="Year")
abline(0, 0)


```



##VIF and multicollinearity
```{r}
library(car)
vif(fit.lm)

```





##Confidence Interval for mean response (mean of Wins)
### for n=1 of data
RankPlayoffs   RankSeason         OOBP           RA           RS         Year 
       5.000        4.000        0.306      600.000      700.000     2012.000 

```{r}
X <- subset(data , select = c(OOBP,RA,RS,Year))
X = as.matrix(X)

x <- as.matrix(X[1,])

MSEreduced = SSEreduced / dfE_reduced

yhat = predict(fit.lm, data=data)

var_hat_Win_hat = t(x) %*% solve(t(X) %*% X) %*% x * MSEreduced

beta <- summary(fit.lm)$coefficients[,1]

x <-rbind(c(1),x)

alpha = 0.05

# CI Upper
CI_Upper <- t(x)%*%beta + qt(alpha/2,dfE_reduced,lower.tail=F) * sqrt(var_hat_Win_hat)

# CI Lower
CI_Lower <- t(x)%*%beta - qt(alpha/2,dfE_reduced,lower.tail=F) * sqrt(var_hat_Win_hat)


print(c(CI_Lower,CI_Upper)) # 92.66599 94.45739

```

##Prediction Interval for mean response (mean of Wins)
### using a new datapoint from 2020 MLB live data
  OOBP           RA           RS         Year 
  0.306      600.000      700.000        2020

```{r}
X <- subset(data , select = c(OOBP,RA,RS,Year))
X = as.matrix(X)

x <- c(0.306, 600.000, 700.000, 2020 )

MSEreduced = SSEreduced / dfE_reduced

yhat = predict(fit.lm, data=data)

var_hat_Win_hat = t(x) %*% solve(t(X) %*% X) %*% x * MSEreduced

beta <- summary(fit.lm)$coefficients[,1]

x <-rbind(c(1),x)

alpha = 0.05

# PI Upper
PI_Upper <- t(x)%*%beta + qt(alpha/2,dfE_reduced,lower.tail=F) * sqrt(var_hat_Win_hat)

# PI Lower
PI_Lower <- t(x)%*%beta - qt(alpha/2,dfE_reduced,lower.tail=F) * sqrt(var_hat_Win_hat)


print(c(CI_Lower,CI_Upper)) # 92.66599 94.45739

```







